---
title: 'Statistical inference assignment: part 1'
author: "MJM Beuken"
date: "5 november 2017"
output: pdf_document
---

## Preparing data, setting parameters and loading packages.

```{r}
## load package
library(ggplot2)

## set constants
labda <- 0.2
n <- 40
N.sim <- 1000

## set seed for reproducible results
set.seed(12345)


## run simulation in matrix
Exp.dist <- matrix(data=rexp(n*N.sim, labda), nrow=N.sim)
Exp.dist.mean <- data.frame(means=apply(Exp.dist, 1, mean))


## show result
head(Exp.dist, 5)
head(Exp.dist.mean, 5)
```


# Show the sample mean and compare it to the theoretical mean of the distribution and show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

### plot histogram with a mean line.

```{r}
p <- ggplot(Exp.dist.mean, aes(x=means)) + 
        geom_histogram(binwidth=0.1, color="black", fill="white")


## Add mean line
p + geom_vline(aes(xintercept=mean(means)), color="blue", linetype="dashed", size=0.5)
```

## Compare sample mean to theoretical mean
### Setting theoretical parameters and calculating sample parameters.

```{r}
mu <- 1/labda
var.theoretical <- (1/labda^2)/n

## calculating sample mean (x (bar))
x.bar <- mean(Exp.dist.mean$means)

## calculating sample variance s²
var.sample <- var(Exp.dist.mean$means)

## show parameters
mu
var.theoretical
x.bar
var.sample
```

## Compare theoretical mean and variance to sample mean and variance in table 

```{r}
Comparing <- matrix(c(mu, var.theoretical, x.bar, var.sample), ncol=2, byrow=TRUE)
colnames(Comparing) <- c("Mean", "Variance")
rownames(Comparing) <- c("Theoretical", "Sample")
Comparing <- as.table(Comparing)
Comparing
```

# Show that the distribution is approximately normal.
## Histogram with normal distribution to get a first graphical indication of normality of sample distribution.

```{r}
ggplot(Exp.dist.mean, aes(x=means)) + 
        geom_histogram(aes(y=..density..), colour="black", fill="white")+
        stat_function(fun=dnorm, color="red", args=list(mean=x.bar), sd=sqrt(var.sample))

## Checking also with qqplot
library("ggpubr")
ggqqplot(Exp.dist.mean$means) + ggtitle("QQplot")
```
###As all the points lay approximately along the reference line, we can assume normality.

## Using Kolmogorov-Smirnov goodness of fit test to compare the sample distribution to the normal distribution based on theoretical parameters.

### Hypothesis under alpha = 5%:
### Ho: The sample data are not significantly different than a normal population.
### Ha: The sample data are significantly different than a normal population.

```{r}
ks.test(Exp.dist.mean, pnorm, mu, sqrt(var.theoretical))
```

### p-value = 0.05122 > 0.025; which means that Ho is failed to be rejected under a 5% probability of a type 1 error.
### The sample distribution isn't significantly different than a normal distribution.

